---
title: "DataFest 2023"
author: "ThreeCurlsAndAGuy"
format: 
  revealjs:
    self-contained: true
editor: visual
---

## Overview

**Aim:** Determine how to target recruitment to improve efficiency.

-   Measure efficiency.

-   Decide on time and expertise focus

**Finding:** Target education experts mid-year due to back-to-school case surge

::: notes

* Assist program --> efficient resource use

* Looked at key areas of inefficiency

* Measured efficiency w/ closure/response time

* Action was recruitment expertise / time

* Found: education inefficient at back-to-school

* Solution: target education experts during this period

:::

## Peak times over the year

```{r Setup}
#Set global chunk options
knitr::opts_chunk$set(
  cache = TRUE
)
#Load packages
library(tidyverse)
library(patchwork)
library(sylcount)
library(lubridate)
library(syuzhet)
library(tidytext)
library(tidyr)

#Read data
questionposts <- read_csv("Data/questionposts.csv")
questions <- read_csv(here::here("Data", "questions.csv"), na = "NULL")
categories <- read_csv(here::here("Data", "categories.csv"))
subcategories <- read_csv(here::here("Data", "subcategories.csv"))

#Record months vector
Months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

#Clean questionposts data
questionposts <- questionposts %>%
  #Get rid of incorrectly expanded columns
  select(
    Id:CreatedUtc
  ) %>%
  #Trim out questionposts items where state abbreviation isn't 2 characters
  filter(str_length(StateAbbr) == 2) %>%
    #Reformat Id as number
  mutate(
    Id = as.numeric(Id)
  )
```

```{r CalculateEfficiency}
questions <- questions %>%
  mutate(
    #Calculate times
    ResponseTime = int_length(interval(ymd_hms(AskedOnUtc), ymd_hms(TakenOnUtc)))/3600,
    ClosureTime = int_length(interval(ymd_hms(TakenOnUtc), ymd_hms(ClosedOnUtc)))/3600,
    Urgency = 1 / int_length(interval(ymd_hms(AskedOnUtc), ymd_hms(LegalDeadline)))/3600,
    Urgency = ifelse(Urgency < 0, NA, Urgency),
    #Calculate when 
    YearAsked = year(AskedOnUtc),
    MonthAsked = month(AskedOnUtc)
  )
```

```{r PeakTimes}
pPeakTimes <- questions %>%
  #Get count of cases in each month & year
  count(Category, MonthAsked, YearAsked) %>%
  filter(YearAsked != 2022) %>%
  #Plot
  ggplot(aes(x = MonthAsked, y = n, color = as.factor(YearAsked))) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  facet_wrap(~Category, scale = "free_y")  + 
  #Format
  scale_x_continuous(breaks = c(1:12), labels = Months)+ 
  theme(axis.text.x = element_text(angle = 90)) + 
  guides(color=guide_legend(title="Year"))

pPeakTimes
```

::: notes

-   Introduce graph: Case count across the year per category

-   Point out: Surge during back to school season for education and juvenile

-   This led us to investigate how well this surge was being handled

-   We will now demonstrate the inefficiency and discuss solutions

:::

# Surge in education

## Affect of surge on education response time

```{r EduResponseByMonth}
pEduResponseByMonth <- questions %>%
  #Get mean response time by month, just for education
  group_by(MonthAsked, Category) %>% 
  summarise(ResponseTimeAvg = mean(ResponseTime, na.rm = TRUE)) %>% 
  filter(Category == "Education") %>%
  #Plot
  ggplot(aes(x = MonthAsked, y = ResponseTimeAvg)) + 
  geom_col() +
  #Format
  scale_x_continuous(breaks = c(1:12), labels = Months) + 
  theme(axis.text.x = element_text(angle = 90)) + 
  guides(fill = guide_legend(title = "Year"))

pEduResponseByMonth
```

::: notes
-   Response time - before the lawyer takes up a case
-   Response time increases at back to school season
:::

## Affect of surge on education closure time

```{r EduClosureByMonth}
pEduClosureByMonth <- questions %>% 
  #Get mean closure time by month, just for education
  group_by(MonthAsked, Category) %>% 
  summarise(ClosureTimeAvg = mean(ClosureTime, na.rm = TRUE)) %>%
  filter(Category == "Education") %>%
  ggplot(aes(x = MonthAsked, y = ClosureTimeAvg)) + 
  geom_col() + 
  #Format
  scale_x_continuous(breaks = c(1:12), labels = Months) + 
  theme(axis.text.x = element_text(angle = 90)) + 
  guides(fill = guide_legend(title = "Year"))

pEduClosureByMonth
```

::: notes
-   Closure time - time lawyer takes up to finish a case
-   Closure time increases at back to school season

-   Very large peak that take up longer time to finish a case
:::

# Flow on effect of education case surge

## Surge effect on response time

```{r AllCatsResponseByMonth}
p3 <- questions %>% 
  group_by(MonthAsked, Category) %>% 
  summarise(ResponseTimeAvg = mean(ResponseTime, na.rm = TRUE)) %>% 
  ggplot(aes(x = MonthAsked, y = ResponseTimeAvg)) + 
  geom_col() + 
  facet_wrap(~Category, scales = "free") +
  #Format
  scale_x_continuous(breaks = c(1:12), labels = Months) + 
  theme(axis.text.x = element_text(angle = 90)) + 
  guides(fill = guide_legend(title = "Year"))

p3
```

::: notes
-   Similar placement of peak in response time increase for juvenile cases
-   Fall and rise in average response time across the board 
-   Mid-year
-   Particularly Work, Employment, Juvenile, Housing
:::

## Surge effect on closure time

```{r AllCatsClosureByMonth}
p2 <- questions %>% 
  group_by(MonthAsked, Category) %>% 
  summarise(ClosureTimeAvg = mean(ClosureTime, na.rm = TRUE)) %>% 
  ggplot(aes(x = MonthAsked, y = ClosureTimeAvg)) + 
  geom_col() + 
  facet_wrap(~Category, scales = "free") +
  #Format
  scale_x_continuous(breaks = c(1:12), labels = Months) + 
  theme(axis.text.x = element_text(angle = 90)) + 
  guides(fill = guide_legend(title = "Year"))

p2
```

::: notes

-   Increase in response time effects closure rates

-   Can be seen in Individual rights and others

-   Therefore a backlog
:::

# Specific surge areas

## Subcategories of education

```{r EduSubcategories}
Time_Education_Questions <- questions %>% 
  filter(Category == "Education") %>%
  mutate(YearAsked = year(AskedOnUtc)) %>% 
  mutate(MonthAsked = month(AskedOnUtc, label = TRUE))

Time_Education_Questions$Subcategory <- ifelse(Time_Education_Questions$Subcategory == "Education/School Discipline/Special Education" |
                                                 Time_Education_Questions$Subcategory == "Special Education" |
                                                 Time_Education_Questions$Subcategory == "School or Special Education" |
                                                 Time_Education_Questions$Subcategory == "Special Education (IEPs, 504 Plans, Behavioral Intervention Plans)" |
                                                 Time_Education_Questions$Subcategory == "Special Education/Learning Disabilities", "Special Education",  Time_Education_Questions$Subcategory)
Time_Education_Questions$Subcategory <- ifelse(Time_Education_Questions$Subcategory == "Education" |
                                                 Time_Education_Questions$Subcategory == "Education Law Cases" |
                                                 Time_Education_Questions$Subcategory == "School/Education" |
                                                 Time_Education_Questions$Subcategory == "School/Juvenile", "School/Education",  Time_Education_Questions$Subcategory)
Time_Education_Questions$Subcategory <- ifelse(Time_Education_Questions$Subcategory == "School Discipline" |
                                                 Time_Education_Questions$Subcategory == "School Discipline (Short-term suspensions, Long-term suspensions, Expulsions)" |
                                                 Time_Education_Questions$Subcategory == "School Discipline (including Expulsion and Suspension)", "School Discipline",  Time_Education_Questions$Subcategory)

Edu_Subcats <- Time_Education_Questions %>%
  group_by(MonthAsked) %>%
  count(Subcategory)

p5 <- ggplot(Edu_Subcats, aes(x = MonthAsked, y = n)) +
  geom_col(aes(fill = MonthAsked)) +
  facet_wrap(~Subcategory)

p5
# Time_Education_Questions <- questions %>% 
#   filter(Category == "Education") %>%
#   mutate(YearAsked = year(AskedOnUtc)) %>% 
#   mutate(MonthAsked = month(AskedOnUtc, label = TRUE))
# 
# Time_Education_Questions$Subcategory <- ifelse(
#   Time_Education_Questions$Subcategory == "Education/School Discipline/Special Education" |
#     Time_Education_Questions$Subcategory == "Special Education" |
#     Time_Education_Questions$Subcategory == "Special Education (IEPs, 504 Plans, Behavioral Intervention Plans)" |
#     Time_Education_Questions$Subcategory == "Special Education/Learning Disabilities", 
#   "Special Education",  
#   Time_Education_Questions$Subcategory
#   )
# 
# Time_Education_Questions$Subcategory <- ifelse(
#   Time_Education_Questions$Subcategory == "Education" | 
#     Time_Education_Questions$Subcategory == "Education Law Cases" | 
#     Time_Education_Questions$Subcategory == "School or Special Education" |
#     Time_Education_Questions$Subcategory == "School/Education" |
#     Time_Education_Questions$Subcategory == "School/Juvenile", 
#   "School/Education/SpecialEd",
#   Time_Education_Questions$Subcategory
#   )
# 
# Time_Education_Questions$Subcategory <- ifelse(
#   Time_Education_Questions$Subcategory == "School Discipline" |
#     Time_Education_Questions$Subcategory == "School Discipline (Short-term suspensions, Long-term suspensions, Expulsions)" |
#     Time_Education_Questions$Subcategory == "School Discipline (including Expulsion and Suspension)", 
#   "School Discipline",  
#   Time_Education_Questions$Subcategory
#   )
# 
# Edu_Subcats <- Time_Education_Questions %>%
#   group_by(MonthAsked) %>%
#   count(Subcategory)
# 
# g4 <- ggplot(Edu_Subcats, aes(x = MonthAsked, y = n)) +
#   geom_col(aes(fill = MonthAsked)) +
#   facet_wrap(~Subcategory)
# 
# g4
```

::: notes
-   Introduce graph: Subcategories over 12 months

-   School discipline starts to rise when school returns, peaks Oct

-   1 in 4 education forum posts discuss suspension/expulsion

-   Lawyers should anticipate and prepare for suspension / expulsion policies
:::

# Consequences

## Sentiment analysis of education queries

```{r}
educationposts <- questions %>%
  left_join(questionposts, by = "QuestionUno") %>%
  select(QuestionUno, PostText, AskedOnUtc, Category) %>%
  filter(Category == "Education") %>%
  mutate(Month = month(AskedOnUtc, label = TRUE))

tidy_eduposts <- educationposts %>%
  unnest_tokens(word, PostText)

cleaned_eduposts <- tidy_eduposts %>%
  anti_join(get_stopwords()) %>%
  arrange(by_group = Month)

bing <- get_sentiments("bing")

edu_sentiment <- cleaned_eduposts %>%
  inner_join(bing) %>%
  count(Month, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative) %>%
  ggplot(aes(Month, sentiment)) +
  geom_bar(aes(fill = Month), stat = "identity", show.legend = FALSE)

edu_sentiment
```

::: notes
-   Introduce graph: Sentiment analysis of discussions on education

-   Downward trend as year progresses

-   not being appropriately managed early on, leading to neg attitude
:::

# Conclusion

## Recommendation

**1. Recruit those with expertise in education shortly before back to school season**

-   This will enable a better response to the case surge in this period

-   Handling that early on will temper increased wait times in several categories

**2. Provide guide materials on Suspension & Expulsion policies**

-   This will decrease closure time in the subcategory of education with the strongest surge
